\LARGE \bf On Physics-Based 3D Waypoint Generation for Autonomous Exploration in Mobile Robotics

### 2011-09-08 paper/On Physics-Based 3D Waypoint Generation/paper.tex
jaja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1106 f4953187-ecf7-0310-90f3-8d0fedb0aae5

In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based data-structure. Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms. Besides its position, each point in the octree also stores its brightness (i.e. intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor. As the platform's measured orientation is less precise that its measured position, this vector's length will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.

@@

We initially planned to use one laserscanner with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig. \ref{lidar_setup_pair}). Besides the obvious disadvantages of another 250 grams of payload and $\sim$10W of power consumption, the platform would have had to constantly pitch during flight to ensure reliable obstacle scanning and detection. To avoid these drawbacks, we instead mounted a single laserscanner as depicted in Fig. \ref{lidar_setup_single}. By constantly yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the platform does not move further than the laserscanners range within one rotation.
% TODO: speed < lidar-range * drehgeschwindigkeit ?

@@

\begin{figure*}[ht]
  \centering
    \includegraphics[width=0.45\textwidth]{images/expl1}
    \includegraphics[width=0.45\textwidth]{images/expl2}
    \includegraphics[width=0.45\textwidth]{images/expl3}
    \includegraphics[width=0.45\textwidth]{images/process5}
    \caption{Mapping setup; bounding volume is blue, detection volume is red, vehicle position is grey, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red. Second figure shows sampling geometry interacting with pointcloud. Geometry hitting the red volume will be converted to a waypoint at the position it last hit a point from the pointcloud. The last figure shows generated pointcloud and leftover sampling geometry.}
    \label{explanation_of_process}
\end{figure*}

As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data. Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input from sensors. Finding the next best view in such data can be very hard, as it does not supply any information about geometric structures such as corners, edges, surfaces and normals. Fortunately, our algorithm does not require such information.

Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped. All generated viewpoints will be constrained to this volume, ensuring that the UAV does not leave the confined space. At initialization, a physics engine is set up with a detection volume below the defined bounding volume. Each point in the pointcloud is automatically registered as a collision object in the physics world. After the octree is populated with an initial set of points (i.e. right after lift-off), the waypoint-generation algorithm is executed as depicted in Fig. \ref{explanation_of_process}:

@@

\begin{figure*}[ht]
  \centering
    \includegraphics[width=0.48\textwidth]{images/process1}
    \includegraphics[width=0.48\textwidth]{images/process2}
    \includegraphics[width=0.48\textwidth]{images/process3}
    \includegraphics[width=0.48\textwidth]{images/process4}
    %\includegraphics[width=0.2\textwidth]{images/process5}
    \caption{Multiple iterations of waypoint generation, as seen from the top, lead to continuous scanning at the border between known and unknown environment.}
    %\label{expl3}
\end{figure*}

@@

>>>>

In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based \cite{meagher1982} data-structure. Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms. Besides its position, each point in the octree also stores its brightness (i.e. intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor. As the platform's measured orientation is less precise that its measured position, this vector's length will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.

@@

We initially planned to use two laserscanners, one with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig. \ref{lidar_setup_pair}). Besides the obvious disadvantages of another 250 grams of payload and $\sim$10W of power consumption, the platform would have had to constantly pitch during flight to ensure reliable obstacle scanning and detection. To avoid these drawbacks, we instead mounted a single laserscanner as depicted in Fig. \ref{lidar_setup_single}. By constantly yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the world is static and the platform does not move further than the laserscanners range within one rotation.
% TODO: speed < lidar-range * drehgeschwindigkeit ?

@@

\begin{figure*}[ht]
  \centering
    \includegraphics[width=0.45\textwidth]{images/process_side/1}
    \includegraphics[width=0.45\textwidth]{images/process_side/2}
    \includegraphics[width=0.45\textwidth]{images/process_side/3}
    \includegraphics[width=0.45\textwidth]{images/process_side/4}
    \caption{Mapping setup; bounding volume is blue, detection volume is red, vehicle position is grey, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red. Second figure shows sampling geometry interacting with pointcloud. Geometry hitting the red volume will be converted to a waypoint at the position it last hit a point from the pointcloud. The last figure shows generated pointcloud and leftover sampling geometry.}
    \label{explanation_of_process}
\end{figure*}

Three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, therefore unorganized point clouds are a very common type of sensor data. Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input from sensors. Finding the next best view in such data can be very hard, as it does not supply any information about geometric structures such as corners, edges, surfaces and normals. Fortunately, our algorithm does not require such information.

Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped. All generated viewpoints will be constrained to this volume, ensuring that the vehicle does not leave the confined space. At initialization, a physics engine is set up with a detection volume below the defined bounding volume. Each point in the pointcloud is automatically registered as a collision object in the physics world. After the octree is populated with an initial set of points (i.e. right after lift-off), the waypoint-generation algorithm is executed as depicted in Fig. \ref{explanation_of_process}:

@@

\begin{figure*}[ht]
  \centering
    \includegraphics[width=0.45\textwidth]{images/process_top/1}
    \includegraphics[width=0.45\textwidth]{images/process_top/2}
    \includegraphics[width=0.45\textwidth]{images/process_top/3}
    \includegraphics[width=0.45\textwidth]{images/process_top/4}
    \includegraphics[width=0.45\textwidth]{images/process_top/5}
    \includegraphics[width=0.45\textwidth]{images/process_top/6}
    \caption{Multiple iterations of waypoint generation, as seen from the top, lead to continuous scanning at the border between known and unknown environment.}
    %\label{expl3}
\end{figure*}

@@

% Octrees
\bibitem{meagher1982}
Donald Meagher, Geometric modeling using octree encoding, {\it Computer Graphics and Image Processing}, 19, 2, pp. 129-147, (1982)
